Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2018-08-29 documentation/paper.tex
Started adding the report for ICN grid search
Added single example row for best run
Added table and updated one graph for new ICN grid search; gravitational analysis is in progress

@@

Gravitational differentiation is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

@@

Iterative cluster nucleation is a second original semi-supervised learning algorithm developed as part of this study. It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

@@

During a grid search of these parameters (in addition to regularization parameters of the network), the best accuracy achieved for replication of AP was exactly 100\%. The mean of all epochs within that training run was 98\%.

As seen in Figure \ref{icn_grid_search}, the network learned to replicate AP very accurately. Interestingly, the examples with AP values between the neutron and alpha particle ranges were also classified as such by the network. This may imply the indecisiveness on these examples is fundamental to the data rather than an implementation problem with AP.

@@

@@

Compared to gravitational differentiation, there is a larger delta between the number of errors on the validation set depending on the number of initial training examples used. Those with 128 examples in $\varsigma$ make 85\% more errors on average. However, iterative cluster nucleation with 128 examples still makes 19\% fewer errors than gravitational differentiation with 256 examples in $\varsigma$.

@@

@@

The following table directly compares performance of the semi-supervised learning systems, with the existing PICO-60 neural network provided as a baseline once again. Note that ``Max AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.

\begin{tabular}{|l|l|l|l|}
    \hline
    Technique & Max AP Accuracy & Max Mean by Run & Class-Wise SD \\
    \hline
    Gravitational Differentiation (256 $\varsigma$) & 98\% & 96\% & TODO \\
    \hline
    Gravitational Differentiation (128 $\varsigma$) & TODO & TODO & TODO \\
    \hline
    Iterative Cluster Nucleation (256 $\varsigma$) & 100\% & 98\% & TODO \\
    \hline
    Iterative Cluster Nucleation (128 $\varsigma$) & TODO & TODO & TODO \\
    \hline
    Existing NN TODO & X & X & X \\
    \hline
    AP TODO & X & X & X \\
    \hline
\end{tabular}

>>>>

To ensure that accuracy statistics are as consistent and reliable as possible, every configuration was trained three times, with a different randomly selected validation set each time. Accuracy, precision, recall, and class-wise standard deviation statistics are the mean over those 3 tests.

@@

Gravitational differentiation (GD) is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

@@

Iterative cluster nucleation (ICN) is a second original semi-supervised learning algorithm developed as part of this study. It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

@@

During a grid search of these parameters (in addition to regularization parameters of the network), the best mean accuracy over three runs of an individual configuration was 99\%, using 256 initial labeled examples in $\varsigma$. With 128 examples, the maximum was 98\%; with 64 it was 97\%, and with 32 95\%. From this, it is evident that it is possible to achieve very high accuracy using this technique. It is also clear that, in general, larger initial sets produce higher performance.

The network's highly accurate and decisive predictions can be observed in Figure \ref{icn_grid_search}.

@@

THIS NEEDS TO BE EDITED VVVV

@@

@@

The following table represents the performance of a variety of iterative cluster nucleation configurations tested during the grid search. As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of iterative cluster nucleation configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0 & 0.25 & 0.01 & 1.025 & 80\% & 79\% & 99\% & 0.60 \\
            \hline
            32 & 0 & 0 & 0.01 & 1.025 & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & 0.001 & 0 & 0.01 & 1.05 & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            64 & 0 & 0.25 & 0.02 & 1.025 & 97\% & 95\% & 100\% & 0.27 \\
            \hline
            128 & 0 & 0.5 & 0.01 & 1.05 & 92\% & 90\% & 100\% & 0.35 \\
            \hline
            128 & 0 & 0.5 & 0.02 & 1.05 & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & 0 & 0.5 & 0.01 & 1.05 & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            256 & 0.003 & 0.5 & 0.01 & 1.05 & 98\% & 98\% & 98\% & 0.26 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

The following table directly compares performance of certain successful semi-supervised learning systems, with AP provided as a baseline for comparison. Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Summary of semi-supervised learning systems.}
        \begin{tabular}{|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & Technique & Mean AP Accuracy & Mean Precision & Mean Recall & Mean CWSD \\
            \hline
            32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            $\infty$ & AP & 100\% & 100\% & 100\% & TODO \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}
